---
title: "Analysis of MUSIC output"
output:
  html_notebook: default
  pdf_document: default
---
Analysis of the thor output in R for ChIP-seq data for p300 on human cells at times 0 and 30. There were two replicates at time 30. Data was aligned to human genome hg38 using BWA and subsequently filtered (see *process_alignments.sh* in *~/Documents/Aim_2/Ramos_2010/p300*). Only looking at chromosome 21 for now.   

Version and full command used can be found in file: 


Download the neccessary packages:
```{r}
library(tidyverse)
library(readr)

library(GenomicRanges)
library(rtracklayer)

library(Gviz)
library(biomaRt)

```

Download the data:

```{r}
    chr_summits=tibble()
 

 tmp=as.tibble(read_table2("~/Documents/Aim_2/homo_sapiens_neutrophil/MUSIC_analysis/default_ERs_MUSIC.bed", 
    col_names = FALSE)) 
  colnames(tmp)=c("chr", "start", "end", "name", "log_10_qvalue", "strand", "summit_position", "mappable_trough_position", "fold_change")
  tmp=dplyr::select(tmp,chr,start,	end	,log_10_qvalue, fold_change)
    chr_summits=rbind(chr_summits, tmp)

tail(chr_summits)

music=dplyr::rename(chr_summits, Chrom=chr, Start=start, End=end, Score=log_10_qvalue)

```


```{r}

music_event=GRanges(
  seqnames=paste0("chr",Rle(chr_summits$chr)),
  ranges = IRanges(chr_summits$start, end=chr_summits$end, names=seq(length(chr_summits$start))),
  strand=Rle(strand('*'), c(nrow(chr_summits))),
  score=chr_summits$log_10_qvalue)


```



# Visualisation with Gviz

I loaded the Gviz library and the biomaRt library. 

I created a biomaRt object from ensembl containing the annotations for the human genome. I made sure that I had the correct version of the human genome (hg18). 

This was used to generated gene model data for the region of interest on chromosome 21.


# Genes overlapping with enriched regions

I create a list of genes with attributes that are found in chromosome 21. 

```{r}


load("~/Documents/Aim_2/homo_sapiens_neutrophil/mar2016egs")
 x=replace(egs$chromosome_name, list=which(egs$chromosome_name=="MT"), values=c("M"))
 x["M"]
# 
# 
 egs$chromosome_name=x

filter(music, Chrom=="Mt")
```

And converted this to a Granges object. 

```{r}
gene_regions = 
  GRanges(seqnames = paste0("chr",Rle(egs$chromosome_name )),
          ranges = IRanges( start = egs$start_position,
                            end = egs$end_position ),
          strand = Rle( rep("*", nrow(egs)) ),
          gene = egs$ensembl_gene_id,
          entrez=egs$entrezgene)
head(gene_regions)
```


# Creating gene/promoter lists for comparison with ranciati code

I want to create a list to compare to the list generated from the macs2 peaks and similar programs. 

I decided to be as broad as possible, and will include a regions $\plus 200$bp around the start of the gene to cover the promoter region, as well as the rest of the gene. 

```{r}

egs$TSS = ifelse( egs$strand == "1", egs$start_position, egs$end_position )

#ifelse( egs$stand == "1", egs$TSS-200, egs$start_position)

head(egs)
gene_regions
gene_promoter_regions = 
  GRanges(seqnames = Rle( paste0("chr",egs$chromosome_name) ),
          ranges = IRanges( 
            start=ifelse( egs$strand == "1", egs$TSS-200, egs$start_position),          
            end=ifelse(egs$strand == "-1", egs$TSS+200, egs$end_position)),
          strand = Rle( rep("*", nrow(egs)) ),
          gene = egs$ensembl_gene_id)
gene_promoter_regions
```

Get 3 gene lists, overlaps only in T0, overlaps only in T30, overlaps in both. Removed multiple hits on the same gene. 


```{r}
ovlp3_gene_promoter = findOverlaps(gene_promoter_regions, music_event )

genes_promoter=as.tibble(egs[ ( queryHits( ovlp3_gene_promoter ) ),]) %>%
    add_column(score=music_event[to(ovlp3_gene_promoter)]$score) %>% 
  group_by(entrezgene) %>%
  filter(score==min(score)) %>%
  distinct()


hist(as.numeric(genes_promoter$chromosome_name))

```



Save as a list

```{r}

music_gene_lists=list(genes_promoter=genes_promoter)

save(music_gene_lists, file="~/Documents/Aim_2/homo_sapiens_neutrophil/MUSIC_analysis/music_coding_gene_lists_mar2016")

```


# Changing Thresholds



```{r}


music_arranged=arrange(music, (Score)) %>%
  mutate(qval=10^Score) %>%
  dplyr::rename(Stop=End)
# macs_all_pvalues=tibble(Start=macs_peaks_all_pvalues$start,Stop=macs_peaks_all_pvalues$end, Chrom=macs_peaks_all_pvalues$chr, log10pval=macs_peaks_all_pvalues$`-log10(pvalue)`)
plot(music_arranged$qval)
# boxplot(c(macs$Stop-macs$Start, (macs_all_pvalues$Stop-(macs_all_pvalues$Start))))

    music_top_10000=music_arranged[1:(nrow(music_arranged)/1.9),]
    music_top_5000=music_arranged[1:(nrow(music_arranged)/5.1),]
    music_top_2000=music_arranged[1:(nrow(music_arranged)/12.7),]
    music_top_1000=music_arranged[1:(nrow(music_arranged)/24.7),]

get_threshold_gene_lists=function(macs_arranged, n_500, n_1000, n_2000, n_4000, n_10000){
    macs_top_10000=macs_arranged[1:(nrow(macs_arranged)/n_10000),]
    macs_top_5000=macs_arranged[1:(nrow(macs_arranged)/n_4000),]
    macs_top_2000=macs_arranged[1:(nrow(macs_arranged)/n_2000),]
    macs_top_1000=macs_arranged[1:(nrow(macs_arranged)/n_1000),]
    macs_top_500=macs_arranged[1:(nrow(macs_arranged)/n_500),]
     
  
    macs_top_10000_event=GRanges(
      seqnames=Rle(paste0("chr",macs_top_10000$Chrom)),
      ranges = IRanges(macs_top_10000$Start, end=macs_top_10000$Stop, names=seq(length(macs_top_10000$Start))),
      strand=Rle(strand('*'), c(nrow(macs_top_10000))),
      Score=macs_top_10000$Score)
    
    
    macs_top_5000_event=GRanges(
      seqnames=Rle(paste0("chr",macs_top_5000$Chrom)),
      ranges = IRanges(macs_top_5000$Start, end=macs_top_5000$Stop, names=seq(length(macs_top_5000$Start))),
      strand=Rle(strand('*'), c(nrow(macs_top_5000))),
      Score=macs_top_5000$Score)
    
    
    macs_top_2000_event=GRanges(
      seqnames=Rle(paste0("chr",macs_top_2000$Chrom)),
      ranges = IRanges(macs_top_2000$Start, end=macs_top_2000$Stop, names=seq(length(macs_top_2000$Start))),
      strand=Rle(strand('*'), c(nrow(macs_top_2000))),
      Score=macs_top_2000$Score)
    
    
    macs_top_1000_event=GRanges(
      seqnames=Rle(paste0("chr",macs_top_1000$Chrom)),
      ranges = IRanges(macs_top_1000$Start, end=macs_top_1000$Stop, names=seq(length(macs_top_1000$Start))),
      strand=Rle(strand('*'), c(nrow(macs_top_1000))),
      Score=macs_top_1000$Score)
  
    macs_top_500_event=GRanges(
      seqnames=Rle(paste0("chr",macs_top_500$Chrom)),
      ranges = IRanges(macs_top_500$Start, end=macs_top_500$Stop, names=seq(length(macs_top_500$Start))),
      strand=Rle(strand('*'), c(nrow(macs_top_500))),
      Score=macs_top_500$Score)
  
      
    ovlp3_gene_promoter_10000 = findOverlaps(gene_promoter_regions, macs_top_10000_event )
    ovlp3_gene_promoter_5000 = findOverlaps(gene_promoter_regions, macs_top_5000_event )
    ovlp3_gene_promoter_2000 = findOverlaps(gene_promoter_regions, macs_top_2000_event )
    ovlp3_gene_promoter_1000 = findOverlaps(gene_promoter_regions, macs_top_1000_event )
    ovlp3_gene_promoter_500 = findOverlaps(gene_promoter_regions, macs_top_500_event )
  
    #T0_gene_promoter=gene_promoter_regions[to(ovlp3_T0_gene_promoter)]
    #T30_gene_promoter=gene_promoter_regions[to(ovlp3_T30_gene_promoter)]
    

      genes_promoter_500=as.tibble(egs[ ( queryHits( ovlp3_gene_promoter_500 ) ),]) %>%
      add_column(Score=macs_top_500_event[to(ovlp3_gene_promoter_500)]$Score) %>% 
      group_by(entrezgene) %>%
      filter(Score==max(Score))   %>%
      distinct() %>%
      ungroup() %>%
      arrange(Score)
        
    genes_promoter_1000=as.tibble(egs[ ( queryHits( ovlp3_gene_promoter_1000 ) ),]) %>%
        add_column(Score=macs_top_1000_event[to(ovlp3_gene_promoter_1000)]$Score) %>% 
      group_by(entrezgene) %>%
      filter(Score==max(Score))   %>%
      distinct() %>%
      ungroup() %>%
      arrange(Score)
    
    
    genes_promoter_2000=as.tibble(egs[ ( queryHits( ovlp3_gene_promoter_2000 ) ),]) %>%
        add_column(Score=macs_top_2000_event[to(ovlp3_gene_promoter_2000)]$Score) %>% 
      group_by(entrezgene) %>%
      filter(Score==max(Score)) %>%
      distinct() %>%
      ungroup() %>%
      arrange(Score)
    
    genes_promoter_4000=as.tibble(egs[ ( queryHits( ovlp3_gene_promoter_5000 ) ),]) %>%
        add_column(Score=macs_top_5000_event[to(ovlp3_gene_promoter_5000)]$Score) %>% 
      group_by(entrezgene) %>%
      filter(Score==max(Score)) %>%
      distinct() %>%
      ungroup() %>%
      arrange(Score)
    
    genes_promoter_10000=as.tibble(egs[ ( queryHits( ovlp3_gene_promoter_10000 ) ),]) %>%
        add_column(Score=macs_top_10000_event[to(ovlp3_gene_promoter_10000)]$Score) %>% 
      group_by(entrezgene) %>%
      filter(Score==max(Score)) %>%
      distinct() %>%
      ungroup() %>%
      arrange(Score)
    
    return(list(genes_promoter_500, genes_promoter_1000, genes_promoter_2000, genes_promoter_4000, genes_promoter_10000))

}



music_threshold_genes=get_threshold_gene_lists(macs_arranged = music_arranged,n_500=46.7, n_1000 = 24.35, n_2000 = 12.46,n_4000 =  6.2, n_10000 = 1)


```




```{r}

genes_promoter_500=music_threshold_genes[[1]]
genes_promoter_1000=music_threshold_genes[[2]]
genes_promoter_2000=music_threshold_genes[[3]]
genes_promoter_4000=music_threshold_genes[[4]]
genes_promoter_all=music_threshold_genes[[5]]

tail(genes_promoter_1000)
tail(genes_promoter_2000[1:1000,])
tail(genes_promoter_4000[1:1000,])
tail(genes_promoter_all[1:1000,])

plot(genes_promoter_all$Score, col="red", type="l")+ #ylim=c(-70,-10), xlim=c(0,5000))+
  points(genes_promoter_4000$Score, col="blue", type="l")+
  points(genes_promoter_2000$Score, col="green", type="l")+
  points(genes_promoter_1000$Score, col="orange", type="l")+
  points(genes_promoter_500$Score, col="pink", type="l")


upset(fromList(list(top_500=genes_promoter_500$entrezgene, top_1000=genes_promoter_1000$entrezgene, top_2000=genes_promoter_2000$entrezgene, top_5000=genes_promoter_4000$entrezgene, all=genes_promoter_all$entrezgene)))


save(genes_promoter_500, genes_promoter_1000, genes_promoter_2000, genes_promoter_4000, genes_promoter_all, file="~/Documents/Aim_2/homo_sapiens_neutrophil/MUSIC_analysis/music_4_thresholds_genes_mar2016")

```
